\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2020-11-17 manuscript.Rnw
Start on first example model

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
@

@@

%% -- Models with aggregate level data -----------------------------------------
\section{Economic modeling with aggregate level data} \label{sec:modeling-aggregate}

@@

\end{tikzpicture}
\caption{Four state disease progression model}\label{fig:econ-eval-process-hesim}
\end{figure}

\subsection{Cohort model}

>>>>

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine = R, eps = FALSE, keep.source = TRUE, eval = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
@

@@

%% -- Models with aggregate level data -----------------------------------------
\section{Economic modeling with aggregate level data} \label{sec:modeling-aggregate}
Health economic models are commonly developed using aggregate level data taken from multiple sources. In these cases it is helpful to be able input model parameters rather than fitting a statistical model. We illustrate with the 4 state disease progression model displayed in Figure~\ref{fig:four-state-model}. Patients advance between successively more severe disease states. It is possible to move from State 2 back to State 1 but recovery is not possible from State 3. Death can occur from any state.

@@

\end{tikzpicture}
\caption{Four state disease progression model}\label{fig:four-state-model}
\end{figure}

There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2). For illustration, we will consider a target population approximated by two representative patients of age 60 and 70. The 60-year old patient represents $1/3$ of all patients with the disease while the 70-year old patient represents $2/3$. The \code{hesim_data} class is used to store this data.

<<>>=
library("hesim")
library("data.table")
strategies <- data.table(
  strategy_id = 1:3,
  strategy_name = c("SOC", "New 1", "New 2")
)
patients <- data.table(
  patient_id = 1:2,
  grp_id = 1:2,
  age = c(60, 70),
  patient_wt = c(1/3, 2/3)
)
hesim_dat <- hesim_data(
  strategies = strategies,
  patients = patients
)
print(hesim_dat)
@

\subsection{Cohort model}
\subsubsection{Parameterization}
Model parameters related to transition probabilities are obtained from a hypothetical analysis of a natural history cohort. If data is not collected at unevenly spaced intervals, then transitions can be described with transition intensity matrices, $Q$, by fitting a continuous time multi-state model designed for panel data (e.g., with the \proglang{R} package \pkg{msm}) \citep{jackson2011multi}. We assume that regression coefficients and a covariance matrix for $Q$ are available from such an analysis.

<<q_coef, echo = FALSE>>=
trans <- c("1->2", "1->4", "2->1", "2->3", "3->4")
q_coef_int <- log(c(.9, .2, .5, 1.1, 1.5))
q_coef_time <- rep(log(1.2), 5)
q_coef_age <- rep(log(1.05), 5)
q_coef_soc <- data.table(
  trans = rep(trans, 3),
  var = rep(c("intercept", "time[2,Inf)", "age"), each = 5),
  loghr = c(q_coef_int, q_coef_time, q_coef_age)
)
@

<<q_vcov, echo = FALSE>>=
n_vars <- nrow(q_coef_soc)
A <- matrix(runif(n_vars^2) * 2 - 1, ncol = 15)/30
q_vcov_soc <- t(A) %*% A
@

<<q_vcov_print>>=
q_vcov_soc[1:4, 1:4]
@

If many studies are available, then relative treatment effects can be estimated by means of a network meta-analysis. We assume that an NMA was performed to estimate the relative risk of transitions from state 1 to both state 2 and death. Treatment strategy 1 was the reference arm. Estimation is performed in terms of the log relative risk since it is approximately normally distributed.

The utility associated with each state is assumed to follow a beta distribution, which is bounded between 0 and 1. The beta distribution is defined in terms of 2 shape parameters, but can be derived from the mean and standard error using the method of moments. Point estimates for utility in states 1, 2, and 3 are 0.9, 0.75, and 0.5, respectively. Standard errors in the same three states are 0.02, 0.03, and 0.05. Utility is always assumed to be 0 in death states in \pkg{hesim}.

<<params>>=
params <- list(
  q_coef_soc = q_coef_soc,
  q_vcov_soc = q_vcov_soc,
  lrr_est = c(log(.80), log(.71)),
  lrr_se = c(log(.06), log(.08)),
  u_mean = c(.9, .75, .05),
  u_se = c(.02, .03, .05)
)
@
